---
title: "EDS-214 Reproducible Workflows: Hurricane Hugo Effects on Stream Chemistry"
format: html
execute:
  message: false
  warning: false
---

# Background
The course EDS-214 titled workflows and reproducibility, emphasizes a real world  example of data analysis that benefits from workflow organization and reproducible code. This course has chosen Schaefer et al.'s (2000) study analyzing the effects of Hurricane Hugo on the watersheds of the Loquillo Mountain Watershed in Puerto Rico as a case stude. In this repository, figure 3 from Schaefer et al.'s (2000) paper was reproduced. 
```{r}
#| warning: false
#| message: false
#| echo: false

#calling libraries
library(tidyverse)
library(here)
library(janitor)
library(patchwork)
library(zoo)
library(png)

#sourcing the moving average function for analysis of the data 
source(here("R", "moving-average-function.R"))
```


# Data
Data was downloaded from the [Environmental Data Initiative](https://portal.edirepository.org/nis/metadataviewer?packageid=knb-lter-luq.20.4923064), an open source environmental database. 

doi:10.6073/pasta/f31349bebdc304f758718f4798d25458
```{r}
#| echo: false
#read in data
watershed_data <- read_csv(here::here("outputs", "compiled_data.csv"))
```

# Methods
Data was cleaned and wrangled to include only necessary metrics to complete the recreation of figure 3 from Schaffer et. al. The pre-analysis product was a single dataframe containing the values for all of the locations nutrient contents. 

Once cleaned, the data was analysed to calculate 9-week moving averages. A moving average function was created for clarity and reproducibility. Below is an example of the code calling this function and calculating the values for the 9-week moving average for one of the five nutrients being graphed. This code was run for each of the nutrients.
```{r}
#|eval: false
#calculating 9 day moving average using mutate and group by 
mov_avg_k <- watershed_data |> 
  group_by(sample_id) |> #group by sample sites
  #mutating to create a new column
  mutate(mov_avg_k = sapply(sample_date, #sapply to run on all of the data 
                            moving_average, 
                            sample_date = sample_date, 
                            nutrient_conc = k, 
                            win_size_wks = 9)) |> 
  #selecting for just necessary data to create a plot 
  select(sample_id, sample_date, k, mov_avg_k)

```

```{r}
#| echo: false
#read in moving average data sets
mov_avg_k <- read_csv(here::here("outputs", "mov_avg_k.csv"))
mov_avg_no3_n <- read_csv(here::here("outputs", "mov_avg_no3_n.csv"))
mov_avg_mg <- read_csv(here::here("outputs", "mov_avg_mg.csv"))
mov_avg_ca <- read_csv(here::here("outputs", "mov_avg_ca.csv"))
mov_avg_nh4_n <- read_csv(here::here("outputs", "mov_avg_nh4_n.csv"))
```

# Results
NEED TO FIGURE OUT HOW TO MAKE MY GRAPH APPEAR BY JUST CALLING THE PNG OR JPG
Each dataframe for the individual nutrients was plotted and combined into a single graph. The graph below is as close a replica of the original graph from Schaefer et al. (2000) as possible given differences in data. 
```{r}
#| echo: false
#creating a hurricane line
hurricane <- as.Date("1989-09-18")

# ggplot of the moving  average k 
k_plot <- ggplot(data = mov_avg_k, 
                 aes(x = sample_date, 
                     y = mov_avg_k, 
                     linetype = sample_id)) +
  geom_line() +
  #adding vertical hurricane hugo line
  geom_vline(xintercept = hurricane, linetype = "dashed") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()) +
  labs(title = "Hurricane Hugo Effects on Stream Chemistry",
       y = "K mg |-1")

# ggplot of the moving average NO3_N
no3_n_plot <- ggplot(data = mov_avg_no3_n, 
                     aes(x = sample_date, 
                         y = mov_avg_no3_n, 
                         linetype = sample_id)) +
  geom_line() +
  #adding vertical hurricane hugo line
  geom_vline(xintercept = hurricane, linetype = "dashed") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()) +
  labs(y = "NO3-N ug |-1")

# ggplot of the moving average NO3_N
mg_plot <- ggplot(data = mov_avg_mg, 
                  aes(x = sample_date, 
                      y = mov_avg_mg, 
                      linetype = sample_id)) +
  geom_line() +
  #adding vertical hurricane hugo line
  geom_vline(xintercept = hurricane, linetype = "dashed") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()) +
  labs(y = "Mg mg |-1")

# ggplot of the moving average NO3_N
ca_plot <- ggplot(data = mov_avg_ca, 
                  aes(x = sample_date, 
                      y = mov_avg_ca, 
                      linetype = sample_id)) +
  geom_line() +
  #adding vertical hurricane hugo line
  geom_vline(xintercept = hurricane, linetype = "dashed") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()) +
  labs(y = "Ca mg |-1")

# ggplot of the moving average NH4_N
nh4_n_plot <- ggplot(data = mov_avg_nh4_n, 
                     aes(x = sample_date, 
                         y = mov_avg_nh4_n, 
                         linetype = sample_id)) +
  geom_line() +
  #adding vertical hurricane hugo line
  geom_vline(xintercept = hurricane, linetype = "dashed") +
  labs(x = "Years",
       y = "NH4-N ug |-1")

#compiling the data to create one graph with all of the required subgraphs
fig3 <- k_plot / no3_n_plot / mg_plot / ca_plot / nh4_n_plot +
 plot_layout(guides = 'collect')

#saving the the fig3 figure into the figs folder
ggsave(here::here("figs", "fig3.png"))

#displaying figure
fig3

#show graph in quarto doc
#source(here::here("R", "ggplotting.R"))


#ggsave(here::here("figs", "fig3.jpg"))

```

Figure 3. Concentrations of 9-week averages of potassium (k), nitrate-N, magnesium (mg), calcium (ca), and ammonium-N (Nh4-n) in four streams in the Loquillo Mountain Watershed before and after Hurricane Hugo.   The vertical lines indicate Hurricane Hugo, the event of study.